---
title: "TreeThinking"
author: "Soren Pedersen"
date: "2023-01-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Introduction###

This document is based on the PP presentation by Ben Marwick CES2021 at the Cultural Evolution Society Conference, 
Sapporo, June 2021 -> the PP is in the literature folder

Downloaded from: https://github.com/benmarwick/CES2021

#Install packages

```{r, install packages, echo=TRUE, eval=FALSE}
install.packages("here")
install.packages("cluster")
install.packages("ggplot2")
install.packages("readr")
install.packages("Momocs")
install.packages("dplyr")
install.packages("maptree")
install.packages("ggimage")
install.packages("phangorn")
install.packages("rworldmap")
install.packages("raster") 
install.packages("NbClust")

#You need to install ggtree, but this is build under older versions. Therefore download it via Biocmaneger

if (!requireNamespace("ggtree", quietly = TRUE)){
  if (!requireNamespace("BiocManager", quietly = TRUE)){
    install.packages("BiocManager")
    BiocManager::install("ggtree")
    library(ggtree)
  } else {
    BiocManager::install("ggtree")
    library(ggtree)
  }
} else {library(ggtree)}

#Faster is to:
install.packages("BiocManager")
BiocManager::install("ggtree")
```

#Load packages

```{r, load packages, echo=TRUE}
library(here)
library(cluster)
library(ggplot2)
library(readr)
library(Momocs)
library(dplyr)
library(maptree)
library(ggimage)
library(phangorn)
library(rworldmap)
library(raster) 
library(NbClust)

#If ggtree was not loaded above
library(ggtree)
```
###THIS PART IS FROM THE OLD EFA_WISHKERMAN DATA###

### (0) Load coordinate data from DiaOutline ###

#Load data into variable lf

```{r, echo=TRUE}
lf <- list.files(path = here::here("data/DiaOutlineTraces_NewNames"), 
                 pattern = "\\.txt$", # this tells the function to only look for .txt type files
                 full.names=TRUE)  # We have to keep the full names on to read in the outlines later.
```

#Create a new variable and split the data using lf_structure()

```{r, echo=TRUE}
lf1 <- Momocs::lf_structure(lf,  
                    split = "/", # Nick: this specifies the separator for splitting. In your case it is a /
                    trim.extension = FALSE)
```

#Create a variable and substr()

```{r, echo=TRUE}
#(3) The function data.frame() creates data frames, tightly coupled collections of variables which share many of the properties of matrices and of lists

#"In order to specify explanatory variables going along with the coordinate or coefficient set, grouping factors or covariates can be specified through a data.frame(), and then used to create subsets (see ?Coo)." (Bonhomme et al. 2014)

lf2<-data.frame(substr(lf1$V7, 1, 4))

#ORIGINAL CODE: lf2<-data.frame(substr(lf1$V2, 1, 5))

#...and name the data frame column

names(lf2)[1] <- "Type"

#ORIGINAL CODE: names(lf2)[1] <- "Type"
```

#Import the coordinates from the .txt DiaOutline files to a new variable "coordinates"

```{r, echo=TRUE}
coordinates <- import_txt(lf)

#ORIGINAL CODE: coordinates <- import_txt(lf)
```

#Build an Out object

```{r, echo=TRUE}
allDiatomOutlines <- Out(coordinates, fac = lf2)

#ORIGINAL CODE: allDiatomOutlines<-Out(coordinates, fac = lf2)
```

#Save the data in RDS format

```{r, echo=TRUE}
saveRDS(allDiatomOutlines, file = "DOcoordinates.RDS")
```

### (1) Load and prepare data ###
#OBS: Here starts the Warnock code

#Load data

```{r, load data, echo=TRUE}
#OBS: make your own file path
#OBS: look at his dataset before loading

outlines_combined_nicolas_2016 <- readRDS("DOcoordinates.RDS")
```

#Prepare data: coo_center

```{r, PrepData_center, echo=TRUE}
outlines_combined_nicolas_2016_centered <- Momocs::coo_centre(outlines_combined_nicolas_2016)
```

#Prepare data: coo_scale

```{r, PrepData_scale, echo=TRUE}
outlines_combined_nicolas_2016_centered_scaled <- Momocs::coo_scale(outlines_combined_nicolas_2016_centered)
```

#Prepare data: coo_slidedirection

```{r, PrepData_direction, echo=TRUE}
outlines_combined_nicolas_2016_centered_scaled <- Momocs::coo_slidedirection(outlines_combined_nicolas_2016_centered_scaled, 
                             direction = "up")

#One code chunk was left out of the original document:

# Momocs::panel(outlines_combined_nicolas_2016_centered_scaled)
```


### (2) Unification of outlines with catalogue-dataframe ###

#Read in .csv files

```{r, echo=TRUE}
#OBS: Make your own filepath and dataset
#OBS: look at his dataset before loading

nicolas_fleches_2016_catalog_ids_coordinates <- read.table(file = here("data/SFP_catalog_ID_site_date_vessel_TAB.txt"), sep="\t", header=TRUE)

#ORIGINAL CODE: nicolas_fleches_2016_catalog_ids_coordinates <- readr::read_csv(file = here("data/nicolas_fleches_2016_catalog_ids_with_coordinates.csv"))
```

#Set names

```{r, echo=TRUE}
outlines_combined_nicolas_2016_names <- 
  names(outlines_combined_nicolas_2016_centered_scaled)
```

#Split names

```{r, echo=TRUE}
outlines_combined_nicolas_2016_names_splitted <- 
  strsplit(outlines_combined_nicolas_2016_names, split = "_")
```

#Make a list() of ID and artifact ID

```{r, echo=TRUE}
ID_and_artefact_ID_list <- list()
for (name_index in 1:length(outlines_combined_nicolas_2016_names)){
  
  ID_and_artefact_ID_interrim_df <- data.frame(ID = paste0(outlines_combined_nicolas_2016_names_splitted[[name_index]][1], "-", outlines_combined_nicolas_2016_names_splitted[[name_index]][2]),
                                               ID_artefact <- outlines_combined_nicolas_2016_names[[name_index]])
  names(ID_and_artefact_ID_interrim_df) <- c("ID", "ID_artefact")
  ID_and_artefact_ID_list[[name_index]] <- ID_and_artefact_ID_interrim_df
  
}

ID_and_artefact_ID_df <- 
  do.call("rbind", ID_and_artefact_ID_list)

#Joins the data in the first column "ID" 

#nicolas_fleches_2016_catalog_ids_coordinates_artefact_ID <- dplyr::inner_join(ID_and_artefact_ID_df, nicolas_fleches_2016_catalog_ids_coordinates, by = "Region")

#ORIGINAL CODE: nicolas_fleches_2016_catalog_ids_coordinates_artefact_ID <- dplyr::inner_join(ID_and_artefact_ID_df, nicolas_fleches_2016_catalog_ids_coordinates, by = "ID")
```

#Make an Out object

```{r, echo=TRUE}
#OBS: here is the first 'fac'

outlines_combined_nicolas_2016_centered_scaled <- 
  Momocs::Out(outlines_combined_nicolas_2016_centered_scaled$coo,
              fac = ID_and_artefact_ID_df) #OBS: here I entered "ID_and_artifact_ID_df", perhaps not correct?

#ORIGINAL CODE: 
#outlines_combined_nicolas_2016_centered_scaled <- Momocs::Out(outlines_combined_nicolas_2016_centered_scaled$coo, fac = nicolas_fleches_2016_catalog_ids_coordinates_artefact_ID)
```

#Remove fragmented outliers

```{r, echo=TRUE}
#OBS: Perhaps you do not need this step!

outlines_combined_nicolas_2016_centered_scaled <- 
  Momocs::filter(outlines_combined_nicolas_2016_centered_scaled, 
                 !ID_artefact %in% c("UK_60_XX_pseudo_no_10", "UK_15_XX_pseudo_no_4"))
```

#List the number of artifacts

```{r, echo=TRUE}
length(which(nicolas_fleches_2016_catalog_ids_coordinates$Site == "Dia Mara"))
length(which(nicolas_fleches_2016_catalog_ids_coordinates$Site == "Dia Shoma"))
length(which(nicolas_fleches_2016_catalog_ids_coordinates$Site == "Jenne-jeno"))
length(which(nicolas_fleches_2016_catalog_ids_coordinates$Site == "KNTII"))
length(which(nicolas_fleches_2016_catalog_ids_coordinates$Site == "Korounkorokale"))
length(which(nicolas_fleches_2016_catalog_ids_coordinates$Site == "Ounjougou"))
length(which(nicolas_fleches_2016_catalog_ids_coordinates$Site == "Timbuktu"))
length(which(nicolas_fleches_2016_catalog_ids_coordinates$Site == "Toubel"))
length(which(nicolas_fleches_2016_catalog_ids_coordinates$Site == "Winde Koroji"))
length(which(nicolas_fleches_2016_catalog_ids_coordinates$Site == "Dhar Tichitt"))
length(which(nicolas_fleches_2016_catalog_ids_coordinates$Site == "Karkarichinkat"))

#ORIGINAL CODE BELOW
#length(which(nicolas_fleches_2016_catalog_ids_coordinates_artefact_ID$country == "Denmark")) # number of artefacts from denmark
#length(which(nicolas_fleches_2016_catalog_ids_coordinates_artefact_ID$country == "France")) # number of artefacts from france
#length(which(nicolas_fleches_2016_catalog_ids_coordinates_artefact_ID$country == "United Kingdom")) # number of artefacts from the uk
```

#Calibrate harmonics

```{r, echo=TRUE}
# Estimates the number of harmonics required for the Fourier methods implemented in Momocs

outlines_combined_nicolas_2016_centered_scaled_harmonics <- 
  Momocs::calibrate_harmonicpower_efourier(outlines_combined_nicolas_2016_centered_scaled, 
                                           plot = F)

# outlines_combined_nicolas_2016_centered_scaled_harmonics
```

#efourier

```{r, echo=TRUE}
outlines_combined_nicolas_2016_centered_scaled_efourier <- 
  Momocs::efourier(outlines_combined_nicolas_2016_centered_scaled,
                   nb.h = as.matrix(outlines_combined_nicolas_2016_centered_scaled_harmonics[["minh"]])[[4,1]], # harmonics for 99.9%
                   norm = F) # you selected `norm=TRUE`, which is not recommended. See ?efourier --> probably no problem in our case
```

#PCA

```{r, echo=TRUE}
outlines_combined_nicolas_2016_centered_scaled_PCA <- 
  Momocs::PCA(outlines_combined_nicolas_2016_centered_scaled_efourier) # PCA on Coe objects, using prcomp.
```

#Minimum number of PCs

```{r, echo=TRUE}
minimum_no_of_pcs_nicolas <- 
  Momocs::scree_min(outlines_combined_nicolas_2016_centered_scaled_PCA,
                    prop = 0.95) # minimum number of axis to use to retain a given proportion (i.e. prop = 0.99 to describe 99% of the variation) -- reduces computing time in the phylogeny estimation step:
```

#Make a scree_plot of the PCA

```{r, echo=TRUE}
Momocs::scree_plot(outlines_combined_nicolas_2016_centered_scaled_PCA)
```


### (3) Create groups ###

#pch.group

```{r, echo=TRUE}
pch.group <- c(rep(21, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Dia Mara"))), 
               rep(22, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Dia Shoma"))),
               rep(23, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Jenne-jeno"))),  
               rep(24, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "KNTII"))),  
               rep(25, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Korounkorokale"))),  
               rep(26, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Ounjougou"))),  
               rep(27, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Timbuktu"))),  
               rep(28, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Toubel"))),  
               rep(29, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Winde Koroji"))),  
               rep(30, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Dhar Tichitt"))),  
               rep(31, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Karkarichinkat"))))


#ORIGINAL CODE:
#pch.group <- c(rep(21, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$country == "Denmark"))), rep(22, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$country == "France"))), rep(23, times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$country == "United Kingdom"))))

```

#col.group

```{r, echo=TRUE}

col.group <- c(rep("skyblue2", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Dia Mara"))), 
               rep("gold", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Dia Shoma"))),
               rep("green2", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Jenne-jeno"))), 
               rep("darkblue", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "KNTII"))), 
               rep("red1", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Korounkorokale"))), 
               rep("orange1", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Ounjougou"))), 
               rep("yellow1", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Timbuktu"))), 
               rep("purple", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Toubel"))), 
               rep("magenta", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Winde Koroji"))), 
               rep("grey", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Dhar Tichitt"))), 
               rep("red3", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID == "Karkarichinkat"))))

#ORIGINAL CODE:

#col.group <- c(rep("skyblue2", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$country == "Denmark"))), rep("gold", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$country == "France"))), rep("green2", times=length(which(outlines_combined_nicolas_2016_centered_scaled_PCA$fac$country == "United Kingdom"))))
```

#Plot

```{r, echo=TRUE}
plot(outlines_combined_nicolas_2016_centered_scaled_PCA$x[,1],
     outlines_combined_nicolas_2016_centered_scaled_PCA$x[,2],
     xlab=paste("PCA 1 (", round(summary(outlines_combined_nicolas_2016_centered_scaled_PCA)$importance[2]*100, 1), "%)", sep = ""),
     ylab=paste("PCA 2 (", round(summary(outlines_combined_nicolas_2016_centered_scaled_PCA)$importance[5]*100, 1), "%)", sep = ""),
     pch=pch.group,
     col="black",
     bg=col.group,
     cex=1,
     las=1,
     asp=1,
     main = "SFP 2023; PCA of vessel outline shapes")

# Add grid lines
abline(v=0, lty=2, col="grey50")
abline(h=0, lty=2, col="grey50")

# Add legend
legend("bottomleft", 
       legend=c("Dia Mara", "Dia Shoma", "Jenne-jeno", "KNTII", "Korounkorokale", "Ounjougou", "Timbuktu", "Toubel", "Winde Koroji", "Dhar Tichitt", "Karkarichinkat"), 
       col="black", 
       pt.bg=c("skyblue2", "gold", "green2", "darkblue", "red1", "orange1", "yellow1", "purple", "magenta", "grey", "red3"), 
       pch=c(21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31), 
       pt.cex=1.5)
```


### (4) Analysis of the outliers ###

#Remove outliers

```{r, echo=TRUE}
nicolas_outliers_db <- 
  fpc::dbscan(outlines_combined_nicolas_2016_centered_scaled_PCA$x, 
              eps = 0.3, 
              MinPts = 3)
```

#Plot outliers 

```{r, echo=TRUE}
plot(nicolas_outliers_db, 
     outlines_combined_nicolas_2016_centered_scaled_PCA$x, 
     main = "DBSCAN", 
     frame = FALSE)
```

#Make a data frame

```{r, echo=TRUE}
nicolas_outliers_cluster <- 
  data.frame(name = row.names(outlines_combined_nicolas_2016_centered_scaled_PCA$x), 
             value = nicolas_outliers_db$cluster, 
             row.names = NULL)
```

#Subset outliers

```{r, echo=TRUE}
nicolas_outliers_cluster_outlier_names <- 
  subset(nicolas_outliers_cluster, value != 1)
```

#(Dont know)

```{r, echo=TRUE}
outlines_combined_nicolas_2016_centered_scaled_PCA$fac$outlier_names <- NA
outlines_combined_nicolas_2016_centered_scaled$fac$cluster <- as.factor(nicolas_outliers_cluster$value)

for (vector_index in 1:length(match(nicolas_outliers_cluster_outlier_names$name, 
                                    outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID_artefact))){
  
  current_index <- match(nicolas_outliers_cluster_outlier_names$name, 
                         outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID_artefact)[vector_index]
  
  outlines_combined_nicolas_2016_centered_scaled_PCA$fac$outlier_names[current_index] <- nicolas_outliers_cluster_outlier_names$name[vector_index]
}
```

#Look at the shape of the outliers

```{r, echo=TRUE}
nicolas_2016_with_outliers <- 
  Momocs::slice(outlines_combined_nicolas_2016_centered_scaled, 
                match(nicolas_outliers_cluster_outlier_names$name, 
                      outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID_artefact))
```

#Plot outliers

```{r, echo=TRUE}
Momocs::panel(nicolas_2016_with_outliers,
              fac = "country_code",
              names = T,
              col = "grey")
```


### (5) Analysis without outliers ###

#Make a new dataset

```{r, echo=TRUE}
nicolas_2016_without_outliers <- 
  Momocs::slice(outlines_combined_nicolas_2016_centered_scaled, 
                -match(nicolas_outliers_cluster_outlier_names$name, 
                       outlines_combined_nicolas_2016_centered_scaled_PCA$fac$ID_artefact))
```

#Calibrate harmonics

```{r, echo=TRUE}
nicolas_2016_without_outliers_harmonics <- 
  Momocs::calibrate_harmonicpower_efourier(nicolas_2016_without_outliers, 
                                           plot = F)  # Estimates the number of harmonics required for the Fourier methods implemented in Momocs. This is the only step in this section that produces data we need in the subsequent step.
```

#Efourier

```{r, echo=TRUE}
nicolas_2016_without_outliers_efourier <- 
  Momocs::efourier(nicolas_2016_without_outliers,
                   nb.h = as.matrix(nicolas_2016_without_outliers_harmonics[["minh"]])[[4,1]], # harmonics for 99.9%
                   norm = F)
```

#PCA

```{r, echo=TRUE}
nicolas_2016_without_outliers_PCA <- 
  Momocs::PCA(nicolas_2016_without_outliers_efourier) # PCA on Coe objects, using prcomp.
```

#Plot PCA

```{r, echo=TRUE}
nicolas_2016_screeplot_wo_outliers <- 
  Momocs::scree_plot(nicolas_2016_without_outliers_PCA, 
                     nax = 1:50)
```

#See how many PCs will give 99.9% - 39 PCs gets there

```{r, echo=TRUE}
summary(nicolas_2016_without_outliers_PCA)
```

#Insert (39) 99.9% outliers

```{r, echo=TRUE}
minimum_no_of_pcs_nicolas_without_outliers <- 39
  # ncol(nicolas_2016_without_outliers_PCA$x) #use all PC axes (Left out in original)
```

#(Dont know)

```{r, echo=TRUE}
gg_nicolas <- Momocs::PCcontrib(nicolas_2016_without_outliers_PCA,
                                nax = 1:5,
                                sd.r = c(-2,-1,0,1,2))
```

#Plot

```{r, echo=TRUE}
gg_nicolas$gg + 
  geom_polygon(fill="slategrey", col="black") + 
  theme_minimal(base_size = 20) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

#Save data .png

```{r, echo=TRUE}
ggsave(here("figures/nicolas_arrowheads_pca_contrib.png"),
       h = 10, w = 8)
```

#Plot

```{r, echo=TRUE}
#library(ggplot2)
gg_nicolas <- gg_nicolas$gg +
  theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

#Nice PCA plot

#Create pch.group

```{r, echo=TRUE}
pch.group <- c(rep(21, times=length(which(nicolas_2016_without_outliers_PCA$fac$country == "Denmark"))), 
               rep(22, times=length(which(nicolas_2016_without_outliers_PCA$fac$country == "France"))),
               rep(23, times=length(which(nicolas_2016_without_outliers_PCA$fac$country == "United Kingdom"))))
```

#Create col.group

```{r, echo=TRUE}
col.group <- c(rep("skyblue2", times=length(which(nicolas_2016_without_outliers_PCA$fac$country == "Denmark"))), 
               rep("gold", times=length(which(nicolas_2016_without_outliers_PCA$fac$country == "France"))),
               rep("green2", times=length(which(nicolas_2016_without_outliers_PCA$fac$country == "United Kingdom"))))
```

#Plot

```{r, echo=TRUE}
plot(nicolas_2016_without_outliers_PCA$x[,1],
     nicolas_2016_without_outliers_PCA$x[,2],
     xlab=paste("PCA 1 (", round(summary(nicolas_2016_without_outliers_PCA)$importance[2]*100, 1), "%)", sep = ""),
     ylab=paste("PCA 2 (", round(summary(nicolas_2016_without_outliers_PCA)$importance[5]*100, 1), "%)", sep = ""),
     pch=pch.group,
     col="black",
     bg=col.group,
     cex=1,
     las=1,
     asp=1,
     main = "Nicolas 2016; PCA of Bell Beaker arrowhead outline shapes without outliers")
# Add grid lines
abline(v=0, lty=2, col="grey50")
abline(h=0, lty=2, col="grey50")
# Add legend
legend("bottomleft", 
       legend=c("Denmark", "France", "United Kingdom"), 
       col="black", 
       pt.bg=c("skyblue2", "gold", "green2"), 
       pch=c(21, 22, 24), 
       pt.cex=1.5)
```


### (6) NJ typochronology ###

#Read .csv data

```{r, echo=TRUE}
typochronologie_csv <- readr::read_csv(here("data/nicolas_2017_typochronologie.csv"))

typochronologie_csv <- dplyr::distinct(typochronologie_csv, ID_country, .keep_all = T)
```

#Subset country (UK)

```{r, echo=TRUE}
typochronologie_csv_UK <- subset(typochronologie_csv, Country == "UK")
```

#Subset country (FR)

```{r, echo=TRUE}
typochronologie_csv_FR <- subset(typochronologie_csv, Country == "FR")
```

#Make a dataframe

```{r, echo=TRUE}
nicolas_2016_without_outliers_PCA_as_df <- 
  as.data.frame(nicolas_2016_without_outliers_PCA$x[,1:minimum_no_of_pcs_nicolas_without_outliers])
```

#Sapply and split dataframe

```{r, echo=TRUE}
nicolas_2016_without_outliers_PCA_as_df$ID_country <- sapply(rownames(nicolas_2016_without_outliers_PCA_as_df),
                                                             function(x){
                                                               strsplit(x, split = "(?<=.{5})", perl = TRUE)[[1]][1] # RegEx "(?<=.{5})" means: split into chunks of 5 characters long
                                                             })
```


### FR ###

#Subset

```{r, echo=TRUE}
nicolas_2016_without_outliers_PCA_as_df_subset_typochron_FR <- subset(nicolas_2016_without_outliers_PCA_as_df, ID_country %in% typochronologie_csv_FR$ID_country)
```

#Make a dataframe

```{r, echo=TRUE}
names_artefacts_ID <- data.frame(artefact_ID = rownames(nicolas_2016_without_outliers_PCA_as_df_subset_typochron_FR),
                                 ID_country = nicolas_2016_without_outliers_PCA_as_df_subset_typochron_FR$ID_country)
```

#dplyr

```{r, echo=TRUE}
names_artefacts_ID_and_period <- dplyr::left_join(names_artefacts_ID, typochronologie_csv_FR[,c("ID_country", "Period")], by = "ID_country")
```

#as.factor

```{r, echo=TRUE}
names_artefacts_ID_and_period$Period <- as.factor(names_artefacts_ID_and_period$Period)
```

#Plot

```{r, echo=TRUE}
plot(nicolas_2016_without_outliers_PCA_as_df_subset_typochron_FR$PC1,
     nicolas_2016_without_outliers_PCA_as_df_subset_typochron_FR$PC2,
     xlab="PC1", 
     ylab="PC2",
     cex=1,
     las=1,
     asp=1,
     main = "Nicolas 2016; PCA of French Bell Beaker arrowhead outline shapes without outliers")

# Add grid lines
abline(v=0, lty=2, col="grey50")
abline(h=0, lty=2, col="grey50")
```

#NULL

```{r, echo=TRUE}
nicolas_2016_without_outliers_PCA_as_df_subset_typochron_FR$ID_country <- NULL
```

#dist (euclidean)

```{r, echo=TRUE}
nicolas_typochron_FR_dist <- dist(nicolas_2016_without_outliers_PCA_as_df_subset_typochron_FR, method = "euclidean")
```

#phangorn

```{r, echo=TRUE}
nicolas_typochron_FR_NJ <- phangorn::NJ(nicolas_typochron_FR_dist)
```

#Make additional information data frame "period"

```{r, echo=TRUE}
additional_information_period <- data.frame(Period = names_artefacts_ID_and_period[,c("Period")])
```

#Change rownames

```{r, echo=TRUE}
rownames(additional_information_period) <- names_artefacts_ID_and_period$artefact_ID
```

#Original comment: need to run to at least here before moving to the next script file


### Tree plot ###

#Plot

```{r, echo=TRUE}
NJ_nicolas_ggtree <- ggtree(nicolas_typochron_FR_NJ) %<+% names_artefacts_ID_and_period +
  geom_tiplab(size=3, 
              aes(label = ID_country,
                  color = ID_country)) +
  geom_treescale() + 
  scale_colour_discrete(na.translate = F) + 
  guides(color=guide_legend(title="Site ID"))
```

#View tree

```{r, echo=TRUE}
NJ_nicolas_ggtree
```

